---
title: "Import hydrographic data into MFDB"
output:
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Hydrographic data}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, message=FALSE, echo=FALSE}
library(unittest)
# Redirect ok() output to stderr
options(unittest.output = stderr())
```
```{r, message=FALSE}
library(mfdb)
```

```{r, message=FALSE, echo=FALSE}
# Don't show chatty MFDB messages
logging::setLevel('WARN')
```

Hydrographic products are available from numerous sources. Here, we provide a working example where outputs from a hydrographic model are:

* downloaded from Copernicus Marine Service https://marine.copernicus.eu/
* aggregated over ICES rectangles
* imported into MFDB


# Download monthly hydrographic data
CMEMS Baltic Sea Physical Reanalysis product provides a physical reanalysis for the whole Baltic Sea area from January 1993 and up to minus 1-1.5 year compared to real time. The product is produced by using the ice-ocean model NEMO-Nordic (based on NEMO-3.6, Nucleus for European Modelling of the Ocean) together with a LSEIK data assimilation scheme,
https://data.marine.copernicus.eu/?option=com_csw&view=details&product_id=BALTICSEA_REANALYSIS_PHY_003_011
As a demonstration we retrieve one year of already processed monthly means of bottom temperature (bottomT), but hydrographic products are also available for separate depth layers and at higher spatiotemporal resolution, and can be aggregated before import according to the needs. Note that an account has to be created to download data from Copernicus Marine Service using the script below.

```{r eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
stopifnot(requireNamespace("CFtime", quietly = TRUE)) # Needed to parse dates

# NB: Before running, set
# Sys.setenv(
#   COPERNICUSMARINE_SERVICE_USERNAME="...",
#   COPERNICUSMARINE_SERVICE_PASSWORD="...",
# )

y <- 2018 # year of interest
dat <- CopernicusMarine::cms_download_subset(
  product       = "BALTICSEA_MULTIYEAR_PHY_003_011",
  layer         = "cmems_mod_bal_phy_my_P1M-m",
  variable      = "bottomT",
  region        = c(17, 54, 30, 58.5),     # c(lon_min, lat_min, lon_max, lat_max)
  timerange     = paste0(y, c("-01-01", "-12-31"))
)
```

# Download ICES rectangles

```{r eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
temp <- tempfile()
download.file(url="https://gis.ices.dk/shapefiles/ICES_StatRec_mapto_ICES_Areas.zip", temp)
unzip(temp, exdir="data_provided/ices_rect")
unlink(temp)
```

# Combine ICES rectangles with rnaturalearth::ne_countries() to get sea surface area

```{r eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
ices.rect <- sf::st_read("data_provided/ices_rect/StatRec_map_Areas_Full_20170124.dbf", quiet = T)

## subset ICES rectangles based on hydrographic data extent
ices.rect <- sf::st_crop(ices.rect, dat)

ices.rect.sea <-
  ices.rect |>
  sf::st_difference(
    rnaturalearth::ne_countries(returnclass = "sf", continent="Europe", scale="large") |>
    sf::st_union()
  ) |>
  dplyr::mutate(
    areaSea_km2 = sf::st_area(geometry) |> units::set_units(km^2)
  ) |>
  dplyr::select(
    rect = ICESNAME,
    areaSea_km2 = areaSea_km2
  ) |>
  sf::st_drop_geometry()
```

# Aggregate hydrographic data by ICES rectangles

```{r eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
hydroVar <- aggregate(dat, ices.rect, FUN = mean, na.rm = TRUE) |>
    sf::st_as_sf(long=T) |>
    sf::st_intersection(ices.rect |>
                    dplyr::select(rect=ICESNAME)) |>
    na.omit() |>
    dplyr::left_join(ices.rect.sea) |>
    dplyr::mutate(year = as.numeric(substring(time,1,4)),
                  month = as.numeric(substring(time,6,7)))
```

# Import bottom temperature into mfdb

Import ICES rectangles and import the hydrographic data as a survey index. Then we can retrieve the hydrographic data as a weighted mean using ICES rectangle as a weighting variable:

```{r, eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
mdb <- mfdb(tempfile(fileext = '.duckdb'))

# import ICES rectangles
mfdb_import_area(mdb, data.frame(
    name = as.character(ices.rect.sea$rect),
    size = as.numeric(ices.rect.sea$areaSea_km2)))

# import area size as index for later use
mfdb_import_cs_taxonomy(mdb, "index_type", data.frame(name=c("ices_rect", "bottom_temp")))
mfdb_import_survey_index(mdb,
                         data_source="area_ices_rect",
                         data.frame(index_type = "ices_rect",
                                        year       = hydroVar$year,
                                        month      = hydroVar$month,
                                        areacell   = as.character(hydroVar$rect),
                                        value      = as.numeric(hydroVar$areaSea_km2)))

# import avg bottom temperature by month and ICES rect (under the 'length' field)
mfdb_import_survey_index(mdb,
                   data_source="bottom_temp_CMEMS",
                   data.frame(year       = hydroVar$year,
                              month      = hydroVar$month,
                              areacell   = as.character(hydroVar$rect),
                              value      = as.numeric(hydroVar$bottomT),
                              index_type = "bottom_temp",
                              stringsAsFactors = TRUE))

# extract mean bottom temperature by quarter and area (weighted mean using rectangle area)
dat <- mfdb_survey_index_mean(mdb, c(), list(
                                       ## area = mfdb_unaggregated(),
                                       area = mfdb_group('a1'=c('42G7','42G8','42G9','43G7','43G8','43G9'),
                                                         'a2'=c('44G7','44G8','44G9','45G7','45G8','45G9')),
                                       timestep = mfdb_timestep_quarterly,
                                       year = 2018,
                                       index_type = "bottom_temp",
                                       data_source = 'bottom_temp_CMEMS'),
                              scale_index = 'area_size')[[1]]
dat <- dat[,c('year', 'step', 'area', 'mean')]
dat
```

```{r, echo=FALSE, eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
# NB: Calculated with:
# raw <- hydroVar[hydroVar$year == 2018 & hydroVar$rect %in% c('44G7','44G8','44G9','45G7','45G8','45G9') & hydroVar$month %in% c(1,2,3),]
# sum(raw$bottomT * raw$areaSea_km2) / sum(raw$areaSea_km2)
ok(ut_cmp_equal(dat, read.table(blank.lines.skip = TRUE, header = TRUE, stringsAsFactors = FALSE, text = '
year step area     mean
2018    1   a1 5.065878
2018    1   a2 5.277945
2018    2   a1 5.535443
2018    2   a2 5.409325
2018    3   a1 6.778031
2018    3   a2 6.149436
2018    4   a1 6.090254
2018    4   a2 5.768832
', colClasses = c(NA, 'character', NA, NA)), tolerance = 1e-5), "extract mean bottom temperature by quarter and area")
```

```{r, eval=nchar(Sys.getenv("COPERNICUSMARINE_SERVICE_USERNAME")) > 0}
mfdb_disconnect(mdb)
options(unittest.output = NULL)
```
