---
title: "Storing & grouping other weight measurements"
output:
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Additional weight measurements}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, message=FALSE, echo=FALSE}
logging::setLevel('WARN')
library(unittest)
library(mfdb)

# Empty database & rebuild
if (exists("mdb")) mfdb_disconnect(mdb)
mfdb('vignette-weightmeasurements', destroy_schema = TRUE)

# Remove our attributes from a dataframe - only used for testing
unattr <- function (obj) {
    attributes(obj) <- attributes(obj)[c('names', 'row.names', 'class')]
    obj
}
```

The following examples use the following table_string helper to succintly
define tables:

```{r}
# Convert a string into a data.frame
table_string <- function (text, ...) read.table(
    text = text,
    blank.lines.skip = TRUE,
    header = TRUE,
    stringsAsFactors = FALSE,
    ...)
```

Firstly, connect to a database and set up some areas/divisions:

```{r}
mdb <- mfdb('vignette-weightmeasurements')
mfdb_import_area(mdb, table_string('
name  division size
45G01     divA   10
45G02     divA  200
45G03     divB  400
'))
```

## Importing data

Normally when using ``mfdb_import_survey`` you supply ``count``, ``length``,
``weight`` columns. But other lenghts can also be supplied:

```{r}
mfdb_import_survey(mdb, data_source = "x",
    table_string('
year    month   areacell        species length  age     weight	liver_weight	gonad_weight	stomach_weight
2000    1       45G01           COD     21      2       210	93		82		72
2000    1       45G02           COD     34      3       220	28		93		99
2000    1       45G03           COD     44      4       230	93		85		72
    '))
```

## Querying data

To get any additional weights back when querying, you need to supply
measurements, for example:

```{r}
agg_data <- mfdb_sample_totalweight(mdb, c('age'), list(
    age = mfdb_unaggregated(),
    null = NULL), measurements = c('overall', 'liver', 'gonad', 'stomach'))
agg_data
```
```{r, message=FALSE, echo=FALSE}
stopifnot(ut_cmp_equal(unattr(agg_data[[1]]), table_string('
year step area age total_weight total_liver_weight total_gonad_weight total_stomach_weight
 all  all  all   2          210                 93                 82                   72
 all  all  all   3          220                 28                 93                   99
 all  all  all   4          230                 93                 85                   72
    ')))
```

Single measurements are also possible:

```{r}
agg_data <- mfdb_sample_totalweight(mdb, c('age'), list(
    age = mfdb_unaggregated(),
    null = NULL), measurements = c('stomach'))
agg_data
```
```{r, message=FALSE, echo=FALSE}
stopifnot(ut_cmp_equal(unattr(agg_data[[1]]), table_string('
year step area age total_stomach_weight
 all  all  all   2                   72
 all  all  all   3                   99
 all  all  all   4                   72
    ')))
```

We can also use ``mfdb_sample_meanweight`` to aggregate:

```{r}
agg_data <- mfdb_sample_meanweight(mdb, c('age'), list(
    null = NULL), measurements = c('liver', 'stomach'))
agg_data
```
```{r, message=FALSE, echo=FALSE}
stopifnot(ut_cmp_equal(unattr(agg_data[[1]]), table_string('
year step area age number mean_liver_weight mean_stomach_weight
 all  all  all all      3          71.33333                  81
    '), tolerance = 1e-7))
```

As well as ``mfdb_sample_meanweight_stddev``:

```{r}
agg_data <- mfdb_sample_meanweight_stddev(mdb, c('age'), list(
    null = NULL), measurements = c('overall', 'gonad'))
agg_data
```
```{r, message=FALSE, echo=FALSE}
stopifnot(ut_cmp_equal(unattr(agg_data[[1]]), table_string('
year step area age number mean mean_gonad_weight stddev stddev_gonad_weight
 all  all  all all      3  220          86.66667     10            5.686241
    '), tolerance = 1e-7))
